---
title: "Properties of the Exponential Distribution"
output:
  html_document:
    keep_md: true
    mathjax: default
---
---

# Properties of the Exponential Distribution

This project compares the exponential distribution to the Central Limit Theorum.

We use `rexp(n, lambda)` to simulate the distribution, where `lambda =
0.2` and the mean of the distribution is `1/lambda`. We run 1000
simulations of 40 samples.

In the investigation, we will address the following questions:

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

---
## Setup

```{r, include=F}
library(ggplot2)
```

We begin by setting some variables and creating a data frame of 1000
simulations with 40 samples from the exponential distribution.

```{r, cache = T}
set.seed(1)

lambda = 0.2
samples <- NULL

for (i in 1 : 1000) samples = rbind(samples, rexp(40, rate = lambda))
str(samples)
```

## Comparing the Sample Mean to the Theoretical Mean

1. Show the sample mean and compare it to the theoretical mean of the distribution.

### Sample Mean:

Here we generate the mean for each of the 1000 rows in the data frame
and then the mean of all the means.


```{r, echo=T}
sampleMeans <- apply(samples,FUN = mean, MARGIN=1)
mean(sampleMeans)
```

### Theoretical Mean:

The theoretical mean is given by:

\(1 / lambda \) =

```{r, echo=T}
1/lambda
```

Here we see that the sample mean roughly coincides with the
theoretical mean, just as the Central Limit Theorum predicts.

## Comparing the Sample Variance to the Theoretical Variance

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

### Theoretical Variance

The theoretical variance of the distribution is \((1 / \lambda)^2\):

```{r}
(1/lambda)^2
```

### Sample Variance

Similar to calculating the sample means above, here we calculated the
variance for all 1000 samples of the exponential distribution.

```{r}
sampleVariances <- apply(samples, FUN = var, MARGIN = 1)
mean(sampleVariances)
```

The theoretical variance is then
3. Show that the distribution is approximately normal.

```{r}
g <- ggplot(as.data.frame(sampleMeans), aes(sampleMeans))
g + geom_histogram(binwidth=0.25)
```

## Conclusion
